---
title: "Eric's Code Contributions"
output: html_notebook
---

Started April 12, 2021.

I'll put my code in this document for now.

```{r setup}
library(tidyverse)
library(adegenet)
library(ape)
library(raster)
library(strataG)
library(readxl)
```

# Ort & Pogson 2007 data

Download data from Ort and Pogson from Genbank. Make a vector for accessions listed in paper as EU028349– EU029064.

```{r OrtPogson}

accs <- paste0("EU0",28349:29064)
# this is 716 sequences. Apparently too long for a single entrez query. Will do it as a for loop instead.

for(a in c(1:716)){
  print(paste(a,accs[a]))
  fetched <- entrez_fetch(db = "nuccore", id = accs[a], rettype = "fasta")
  write(fetched, file = "Ort_Pogson_2007.fasta", append = TRUE, sep = "\t")
}

```

Filter out the CO1 data

```{r Filter}

op <- read.FASTA("Ort_Pogson_2007.fasta")
op.co1 <- op[str_detect(names(op), pattern = "cox1")]
names(op.co1) <- str_extract(names(op.co1), pattern = "[A-Z]+[0-9]+[MF]")
write.FASTA(op.co1, file = "Ort_Pogson_CO1.fasta")
```

# Redo IBD analysis with new StrataG

```{r}
myt_aln<-read.FASTA("lat_sorted_seqs.fasta")
myt_aln<-as.matrix(myt_aln)
pop<-gsub(labels(myt_aln),pattern="([A-Za-z_]+)_\\d+",replacement="\\1")
pop<-gsub(pop,pattern="_",replacement="")

poporder<-unique(pop)

myt_g<-sequence2gtypes(myt_aln,strata=pop)
myt_g<-labelHaplotypes(myt_g)
pairwise<-pairwiseTest(myt_g, model = "raw", nrep=0,quietly = F, max.cores=4)
pairwise_phi<-as.dist(pairwiseMatrix(pairwise, stat = "PHIst"))
pairwise_phi <- as.matrix(pairwise_phi)[poporder, poporder]
pairwise_phi <- pairwise_phi/(1-pairwise_phi)
pairwise_phi.d <- as.dist(pairwise_phi)

pairwise_fst <- as.dist(pairwiseMatrix(pairwise, stat = "Fst"))
pairwise_fst <- as.matrix(pairwise_fst)[poporder, poporder]
pairwise_fst <- pairwise_fst/(1-pairwise_fst)
pairwise_fst.d <- as.dist(pairwise_fst)



myt_metadata<-read_xlsx("myt_metadata.xlsx",col_types = c("text","numeric","numeric"))
#myt_unique<-unique(myt_metadata[,c("locality","decimalLatitude","decimalLongitude")]) #pull out just locality and lat/longs, and get unique values
myt_metadata <- myt_metadata %>% arrange(desc(decimalLatitude))



gdist<-pointDistance(p1=myt_metadata[,c(3,2)], lonlat=T,allpairs=T)/1000 #measure great circle distance between points in kilometers

gdist<-as.dist(gdist) #convert to distance object

phi_mantel<-mantel.randtest(pairwise_phi.d,gdist)

phi_plot<- ggplot(tibble(gdist,pairwise_phi.d), aes(x=gdist,y=pairwise_phi.d)) + 
                    geom_point() + 
                    geom_smooth(method=lm) + xlab("Great Circle Distance(km)") +
                    ylab(expression(Phi["ST"]/1-Phi["ST"]))

phi_plot

fst_plot<- ggplot(tibble(gdist,pairwise_fst.d), aes(x=gdist,y=pairwise_fst.d)) + 
                    geom_point() + 
                    geom_smooth(method=lm) + xlab("Great Circle Distance(km)") +
                    ylab(expression(F["ST"]/1-F["ST"]))

fst_mantel<-mantel.randtest(pairwise_fst.d,gdist)


fst_plot



```

# MigraiNe Analysis

Migraine comes with a compilable program `Nexus2GP` that converts haplotype numbers to a genepop format. Could be done in R too, but `Nexus2GP` is probably better to ensure its in the right format.

```{r migraine}

write.nexus.data(myt_aln,file = "Mcalifornianus_210524.nex",interleaved=F)

```

Parmfile for Migraine:

```{bash parmfile1}

GenepopFileName=Mcalifornianus_210524-GP.txt
DemographicModel=LinearIBD
#next two settings describe total size of the habitat (not just the sampled area), 
#https://www.gbif.org/species/2285680
PSONMin=0 0
PSONMax=XXXX 0
# Neighborhood size = 50km
#XXX/50 = GeoBinNbr
GeoBinNbr=17
GeoUnit= ind.km
#alternate way of specifying the habitat, not used for now
#habitatPars= 0.5 0.5 400 1 0
#habitatPars=0 0 0 300 0
#Infinite sites mutation model, will drop sites with more than 2 alleles
MutationModel=ISM

#sampling - could potentially change the sampling parameters, but not right now
#samplingSpace=,,
#samplingScale=,,
#Analysis - this will do 5 iterations of 100 (30 runs/point) points, 
#and overwrite them with 5 runs of 250 points
writeSequence= Over,Over,Over,Over,Over,Append,10
StatisticSequence=PAC
PointNumber=100,100,100,100,100,250
Nrunsperpoint=30,30,30,30,30,50
#Wide priors on Neu, Nem and g 
LowerBound=0.1,1,0
Upperbound=10,10000,1
oneDimCI= 2Nmu, 2Nm, Nb, condS2
CoreNbrForR=4
#Plots= all1DProfiles
1DProfiles=2Nmu, 2Nm, Nb, condS2, g
extrascale=Nb=logscale
graphicFormat=pdf
#writeAdHocFiles=T
```