---
title: "Mytilus californianus full range population genetic analysis"
author: "Vanessa Garcia, Paulina Perez, Alicia Abadia-Cardoso, Eric Crandall"
format:
  html:
    theme: solar
    toc: true
    toc-float: true
    toc-depth: 5
    df-print: paged
    smooth-scroll: true
    citations-hover: true
editor: source
warning: false
code-fold: true
---

# Setup

Load the packages we will need for this analysis

```{r}
library(knitr)
library(ape)
library(pegas)
library(rlang)
library(adegenet)
library(rentrez)
library(tidyverse)
#library(Biostrings)
#library(BiocGenerics)
library(betapart)
library(strataG)
library(readxl)
library(ggmap)
library(raster)
library(phyclust)
library(phangorn)
```

# Introduction

Here we will be conducting a population genetic analysis of the California Mussel (Mejillón de California; *Mytilus californianus*), a species that inhabits the mid-intertidal zone on open coastlines from Baja California to southern Alaska. We have been fortunate to obtain tissue samples and mitochondrial CO1 sequences from the entire range, from near Sitka in Alaska, to Punta Baja in Baja California. We now want to conduct some basic population genetic analyses on these data, and then set up an analysis in Migrate-n , which Eric has shown to be much more sensitive to metapopulation structure than $F_{ST}$-based methods.

# Data Import

The data comprise 501 mtDNA sequences for 480 bp from 23(!) localities, so an average sample size of `r 501/24` per locality. Pretty good! We will import them here from a FASTA file, and split of the locality names, and convert them into formats for various packages.

```{r}
# read into DNAbin format
myt_aln <- read.FASTA("data/mytilus_full_range_230323.fasta") 
# convert from a list DNAbin to a matrix DNAbin, as the sequences were already aligned in Geneious
myt_aln <- as.matrix(myt_aln)
# get the population names and add spaces to them
loc <- gsub(labels(myt_aln),pattern="([A-Za-z_]+)_\\d+",replacement="\\1")
loc <- gsub(loc, pattern="([a-z])([A-Z])", replacement= "\\1 \\2")
#create a list of localities
locorder <- unique(loc)
# convert to gtypes format for StrataG
myt_g <- sequence2gtypes(myt_aln, strata=loc)
# find and label the unique haplotypes
myt_g <- labelHaplotypes(myt_g)

#Warning message:  <-- why is it throwing this warning??!!
#The following samples are missing data for all loci and have been removed: CapeArago_20, #WhitePoint_03, WhitePoint_17 
```
Read in the metadata - just the sampling event metadata for now.

```{r}
events <- read_excel("data/Mcalifornianus_GEOME.xlsx", sheet = "Events")
events <- events %>% mutate(decimalLatitude = as.numeric(decimalLatitude)) %>% 
                    mutate(decimalLongitude = as.numeric(decimalLongitude))

events2 <- events[,c("locality","decimalLatitude", "decimalLongitude")] 
names(events2) <- c("locality", "lat", "lon")

mc_map <- get_map(location = c(top = 60, bottom = 28, left = -145, right = -114 ), zoom = 4)

ggmap(mc_map) + 
  geom_point(data = events2, color = "red", size = 2) +
  geom_text(data = events2, aes(label = paste("  ", as.character(locality), sep="")), 
            angle = 60, color = "yellow", size = 2)

```

# Fst

Calculate pairwise Wright's $F_{ST}$ and $\Phi_{ST}$, which is the sequence analog.

```{r}
pairwise <- pairwiseTest(myt_g, model = "raw", nrep=0, quietly = F, max.cores=4)
pairwise_phi <- pairwiseMatrix(pairwise, stat = "PHIst")
pairwise_phi <- as.matrix(pairwise_phi)
pairwise_phi[upper.tri(pairwise_phi)]<- t(pairwise_phi)[upper.tri(pairwise_phi)]

pairwise_phi <- pairwise_phi[locorder, locorder]
#pairwise_phi <- pairwise_phi/(1-pairwise_phi)
pairwise_phi.d <- as.dist(pairwise_phi)

pairwise_fst <- pairwiseMatrix(pairwise, stat = "Fst")
pairwise_fst <- as.matrix(pairwise_fst)
pairwise_fst[upper.tri(pairwise_fst)]<- t(pairwise_fst)[upper.tri(pairwise_fst)]

pairwise_fst <- pairwise_fst[locorder, locorder]
#pairwise_fst <- pairwise_fst/(1-pairwise_fst)
pairwise_fst.d <- as.dist(pairwise_fst)

heatmap(pairwise_phi)

write.csv(pairwise_phi, "results/pairwise_phi_mcali.csv")
write.csv(pairwise_fst, "results/pairwise_fst_mcali.csv")
```

# Isolation by Distance

```{r}
events2 <- events2 %>% arrange(desc(lat))
gdist <- pointDistance(p1=events2[,c(3,2)], lonlat=T,allpairs=T)/1000 #measure great circle distance between points in kilometers

gdist<-as.dist(gdist) #convert to distance object

phi_mantel<-mantel.randtest(pairwise_phi.d,gdist)

phi_plot<- ggplot(tibble(gdist,pairwise_phi.d), aes(x=gdist,y=pairwise_phi.d)) + 
                    geom_point() + 
                    geom_smooth(method=lm) + xlab("Great Circle Distance(km)") +
                    ylab(expression(Phi["ST"]/1-Phi["ST"]))
phi_plot

ggsave(filename = "phi_plot.pdf", width = 8, height=6,units = "in")


fst_mantel<-mantel.randtest(pairwise_fst.d,gdist)

fst_plot<- ggplot(tibble(gdist,pairwise_fst.d), aes(x=gdist,y=pairwise_fst.d)) + 
                    geom_point() + 
                    geom_smooth(method=lm) + xlab("Great Circle Distance(km)") +
                    ylab(expression(F["ST"]/1-F["ST"]))



fst_plot

ggsave(filename = "Fst_plot.pdf", width = 8, height=6,units = "in")
```

# Haplotype Networks

## Pegas Haplotype Network

Code for haplotype network in pegas

```{r}
#Haplotype Code to visualize genetic variation between populations

d<-dist.dna(myt_aln) # I think this is the right matrix to use
h<-pegas::haplotype(myt_aln)
h<-sort(h,what="label")
net<-pegas::haploNet(h)
#net<-pegas::mst(h)
i<-utils::stack(setNames(attr(h,"index"),rownames(h)))
i<-i[order(i$values),]
ind.hap<-table(hap=i$ind, pop=pop)

ghost.stories<- c("darkslategrey","darkslateblue", "blue4", "royalblue4", "deepskyblue", "royalblue1", "skyblue4","mediumseagreen","darkslateblue","firebrick","coral","darkorange3","darkorange","gold2","gold","yellow1","moccasin","lightskyblue1","lightcyan1")


ind.hap <- ind.hap[,c(11,6,1,13,3,8,16,18,2,5,17,10,15,4,19,12,7,14,9)]
ind.hap


pdf(file = "modified_haplotype.pdf", width = 8.5, height = 11)
plot(net,size=(attr(net,"freq")/5),scale.ratio=2.25,pie=ind.hap,legend=F,labels=F,threshold=0,show.mutation=2,bg=ghost.stories,pch=19,ncol=1,cex=0.85)
legend("topleft",colnames(ind.hap),col=ghost.stories,pch=19,ncol=1,cex=0.85)

dev.off()

```

## PopArt

Code for preparing for PopArt.

```{r}
hapseqs<-h[] # pull out the DNAbin object from the haplotype object #write this to nexus format
write.nexus.data(hapseqs,
                 file="neralb_haplotypes.nex",
interleaved=F) #write both haplotype tables to csv
write.csv(ind.hap, file="neralb_table.csv",
quote=F)
write.csv(ind.hap2, file="neralb_table_IP.csv",
quote=F)
#write the flattened metadata too.
write.csv(event_info, file="neralb_event_metadata.csv",
quote=F)
```


# Migrate

## Preparing the data for migrate

```{r}
# convert to Phangorn
myt_phy <- phyDat(myt_aln)
# create a Jukes-Cantor distance matrix from all sequences
dm <- dist.ml(myt_phy, model ="JC69")
#create a neighbor-joining tree from the distance matrix
tree <- NJ(dm)

mcali_modeltest<-modelTest(myt_phy, tree,G=T, I=F)

mcali_modeltest<-mcali_modeltest[order(mcali_modeltest$BIC),]

head(mcali_modeltest)
# some funky code to get the gamma shape parameter
env <- attr(mcali_modeltest, "env")
TIM1 <- get("TIM1+G(4)", env)
#eval(HKY, env=env)
shape <- 0.26375833217876

plot(density(rgamma(1000,shape=shape)), xlim=c(0,1))
phangorn:::discrete.gamma(shape,4)#shapeparam
migrate_rates <- c(0.002815926, 0.076790475, 0.529615964, 3.390777635)
#TiTvRatio(mcali) #need strataG for this

# pml(tree = `tree_TIM1+G(4)`, data = data, bf = c(0.288481138885885, 
# 0.160432444597162, 0.209795981107601, 0.341290435409352), Q = c(1, 
# 5.39011701149672, 0.297813690823323, 0.297813690823323, 17.1942098908114, 
# 1), model = "TIM1", k = 4, shape = 0.26375833217876)

```