---
title: "Migrate Analysis"
output: html_notebook
---

This is code, originally developed for Crandall et al. 2019, to read through Migrate outputs.

```{r setup}
#install.packages("tidyverse")
library(tidyverse)
library(perm)
library(igraph)
install.packages("network")
install.packages("readxl")
library(network)
library(readxl)


```

## Function for bayes factor calculations
```{r bfcalcs}
# a function for calculating model selection statistics
bfcalcs<-function(df,ml="bezier.corrected"){
  df$thermodynamic<-as.numeric(df$thermodynamic)
  df$bezier.corrected<-as.numeric(df$bezier.corrected)
  df$harmonic<-as.numeric(df$harmonic)
    mlcol<-df[,ml] 
	bmvalue<-mlcol[which.max(mlcol)]
	lbf<-2*(mlcol-bmvalue)
	choice<-rank(-mlcol)
	modelprob<-exp(lbf/2)/sum(exp(lbf/2))
	dfall<-cbind(df,lbf,choice,modelprob)
	return(dfall)
}	
```


## Harvesting Likelihoods

```{r harvest_likelihoods}
#wd<-"~/github/Mytilus/migrate_for_evolution/"
wd<-"~/Mytilus_Project/Code/Mytilus/migrate_for_evolution/"
setwd(wd)
m<-c("Panmixia","Island","Point_Conception","Symmetrical_Stepping_Stone",
     "Stepping_Stone_CC","Stepping_Stone")



likelists<-list()
for(r in 1:20){
  wd1<-paste(wd,"mytilus_",r,sep="")
  setwd(wd1)
  # start here if you just want to do one list
  likelist <- list() #initialize an empty list
  marglike <- data.frame(model=character(0),thermodynamic=numeric(0),
                         bezier.corrected=numeric(0),
                         harmonic.mean=numeric(0),stringsAsFactors=F)  
  l=1 #initialize l
    for(i in m){ #i<-"stepping_stone"
      

      wd3<-file.path(wd1,i)
      print(wd3)
      if(!file.exists(wd3)){next}
      setwd(wd3)
      outfile<-scan(file="outfile.txt",what="character",sep="\n") #scan in the outfile, separating at each newline
      
      #get the result from thermodynamic integration
      thermoline<-grep("(1a)",outfile,value=T) #find the line with the thermodynamic likelihood on it
      if(length(thermoline)==0){next}
      thermoline<-strsplit(thermoline,split="=",fixed=T) #split it up
      thermo<-as.numeric(substr(thermoline[[1]][2],start=1,stop=12)) #grab the thermodynamic likelihood
      bezistart <- grep("\\(",strsplit(thermoline[[1]][2],split="")[[1]])+1
      bezier<-as.numeric(substr(thermoline[[1]][2],start=bezistart,stop=bezistart+11)) #and the bezier-corrected value
      #get the harmonic mean
      harmoline<-grep("\\(2\\) H",outfile,value=T) #find the line with harmonic mean likelihood on it
      harmoline<-strsplit(harmoline,split="=",fixed=T) #grab the harmonic mean
      harmo<-as.numeric(substr(harmoline[[1]][2],start=1,stop=12))
      marglike[l,]<-c(i,thermo,bezier,harmo) #add this as a row to the data frame
      l=l+1
      likelist<-marglike #add the dataframe to the list
    }
  
# stop here if you just want one round  
  likelists[[r]]<-likelist
}
setwd(wd)
```




## Model Selection
On each round, and then bind them into a list, then convert to a dataframe. Take the mean and SD, and then perform model selection on this summarized table. 



```{r model_selection}

modeltable <- map_dfr(likelists, bfcalcs, .id="Replicate")


mytilus_summary <- modeltable %>% group_by(model) %>% summarize(Marg_Like = mean(bezier.corrected), 
                                                     Marg_Like_SD = sd(bezier.corrected)) %>% 
                                                      mutate(Rank = rank(-Marg_Like)) %>% 
                                                      mutate(lbf = 2*(Marg_Like - max(Marg_Like))) %>% 
                                                      mutate(modelprob = exp(lbf/2)/sum(exp(lbf/2)))

                                     
                                        
ttest<-permTS(x=modeltable$bezier.corrected[which(modeltable$model=="Stepping_Stone_CC")],
              y=modeltable$bezier.corrected[which(modeltable$model=="Panmixia")], 
              alternative="greater", method="exact.mc")
 
kmeans.test<-permKS(x = modeltable$bezier.corrected, g = modeltable$model, 
                   method="exact.mc")

write.csv(mytilus_summary,file="BestModeltable_Means_of_20reps.csv")

margLplot <- ggplot(data=modeltable, aes(x=model,y=bezier.corrected)) + geom_violin() +
      geom_point(color="blue",size=0.5) + scale_x_discrete(drop=FALSE) +
      geom_point(data = mytilus_summary, aes(x=model, y=Marg_Like)) +
      theme(axis.text.y = element_text(size=16),legend.position="none",
          axis.title.x=element_text(size=16),axis.title.y=element_blank(),
          plot.title=element_text(size=20)) +
          ylab("Marginal Log-Likelihood") +
          coord_fixed(0.1) #+ coord_flip() apply after gg.gap
  
 ggsave(filename="Migrate_Violins_20reps.pdf", plot = margLplot)


```

#Modify margLplot using discontinuous axis code
```{r}
#if(!require(gg.gap)) install.packages("gg.gap")
#library(gg.gap)

# we need to tweak the theme a bit to make it look nice
# and we need to do it before we pass the plot to gg.gap
gapped.figure <- margLplot +
  theme(
    panel.background = element_rect(fill = "white"), 
    panel.grid = element_blank(),
    axis.line = element_blank()
  )

gapped.figure %>% 
  gg.gap::gg.gap(
    ylim = c(-2300, -2050), 
    segments = list(c(-2250, -2200)),
    tick_width = 150,
    c(0.7,0,0.3)
  )
gapped.figure

final_margL_figure <- gapped.figure + coord_flip()
```


# Make Metapopulation Model Figures

```{r}



#nodes<-read_xlsx("myt_metadata.xlsx",col_types = c("text","numeric","numeric")) %>% 
nodes <- read_csv("myt_metadata.csv") %>%   
        slice(4:15) %>% arrange(desc(decimalLatitude)) %>% mutate(vertex = 1:12)

stepedges<-read.csv("step_edges.csv",header=F)
colrs<-c("black","gray50")

#island model make all possible connections

edges<-t(combn(x=nodes$vertex,m=2))
# and the reverse
edges2<-rbind(edges,edges[,c(2,1)])

#make the n-island model
west_side <- as.matrix(nodes[,c(3,2)])
island<-graph_from_data_frame(edges2,directed=T)

#circle<-layout_in_circle(island,order=c(11,10,9,8,7,6,5,4,3,2,1,14,13,12))


#make a stepping-stone model
stepstone<-graph_from_data_frame(stepedges,directed=T)
#make southward stepping-stone model
cc_stepedges <- stepedges[seq(1,22,2),]
cc_stepstone <- graph_from_data_frame(cc_stepedges, directed=T)

#can't figure out why it won't draw arrows

plot(island,layout=west_side, edge.arrow.size=0.05, edge.curved=0.3, 
     vertex.label=V(island)$locality, 
     vertex.size=8,vertex.label.cex=0.7,vertex.shape="csquare")

plot(stepstone,layout=west_side, edge.arrow.size=0.05, edge.curved=0.3, 
     vertex.label=V(stepstone)$label, vertex.color=V(stepstone)$sector,
     edge.color="black",vertex.label.cex=0.7,vertex.size=8, arrow.mode = 1, arrow.size=1000)

plot(cc_stepstone,layout=west_side, edge.arrow.size=0.05, edge.curved=0.3, 
     vertex.label=V(stepstone)$label, vertex.color=V(stepstone)$sector,
     edge.color="black",vertex.label.cex=0.7,vertex.size=8, arrow.mode = 2, arrow.size=5)





```




